### 2016-06-08 Paper_v4.tex
First p of Results

\subsection*{Reduced order complement network.}
The reduced order complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
Activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
The C3b fragment then recombined with C3 to form the alternate pathway (AP) C3 convertase.
Both the CP and AP C3 convertases catalyzed the cleavage of C3 into C3a and C3b.
A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP (or AP) C5 convertase.
The C5 convertase catalyzed the cleavage of C5 into the C5a and C5b fragments.
Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
Thus, while the reduced order complement model encoded significant biological complexity, it was highly compact consisting of only 18 differential equations and 28 model parameters.
Next, we estimated an ensemble of model parameters from time series measurements of the C3a and C5a complement proteins.

@@

\section*{Materials and Methods}
We used ordinary differential equations (ODEs) to model the time evolution of complement proteins ($x_{i}$) in the reduced order model:
\begin{equation}
	\frac{dx_{i}}{dt}  =  \sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)\qquad{i=1,2,\hdots,\mathcal{M}}\\
\end{equation}
where $\mathcal{R}$ denotes the number of reactions and $\mathcal{M}$ denotes the number of protein species in the model.
The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$.
Typically, reaction $j$ is a non-linear function of biochemical and enzyme species abundance, as well as unknown model parameters $\mathbf{k}$ ($\mathcal{K}\times{1}$).
The quantity $\sigma_{ij}$ denotes the stoichiometric coefficient for species $i$ in reaction $j$.
If $\sigma_{ij}>0$, species $i$ is produced by reaction $j$.
Conversely, if $\sigma_{ij}<0$, species $i$ is consumed by reaction $j$, while $\sigma_{ij} = 0$ indicates species $i$ is not connected with reaction $j$.
Species balances were subject to the initial conditions $\mathbf{x}\left(t_{o}\right) = \mathbf{x}_{o}$.

>>>>

\subsection*{Reduced order complement network.}
The complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
The alternate pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b (not C3w).
The C3b fragment generated by hydrolysis (or by CP C3 convertase) could then form the alternate pathway (AP) C3 convertase (C3bBb).
We did not consider C3w, nor the formation of the initial alternate C3 convertase (C3wBb).
Rather, we assumed C3w was equivalent to C3b and only modeled the formation of the main AP C3 convertase.
Both the CP and AP C3 convertases catalyzed the cleavage of C3 into C3a and C3b.
A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP or AP C5 convertase (C4bC2aC3b or C3bBbC3b).
Both C5 convertases catalyzed the cleavage of C5 into the C5a and C5b fragments.
In this initial study, we simplified the model by assuming both Factor B and Factor D were in excess.
However, we did explicitly account for two control proteins, Factor H and C4BP.
Lastly, we did not consider MAC formation, instead we stopped at C5a and C5b.
Lectin pathway activation, and C3/C5 convertase activity was modeled using a combination of saturation kinetics and non-linear transfer functions, which facilitated a significant reduction in the size of the model while maintaining performance.
Binding interactions were modeled using mass-action kinetics, where we assumed all binding was irreversible.
Thus, while the reduced order complement model encoded significant biology, it was highly compact consisting of only 18 differential equations and 28 model parameters.
Next, we estimated an ensemble of model parameters from time series measurements of the C3a and C5a complement proteins.

@@

\section*{Materials and Methods}

\subsection*{Formulation and solution of the complement model equations.}
We used ordinary differential equations (ODEs) to model the time evolution of complement proteins ($x_{i}$) in the reduced order model:
\begin{equation}
	\frac{dx_{i}}{dt}  =  \sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)\qquad{i=1,2,\hdots,\mathcal{M}}\\
\end{equation}
where $\mathcal{R}$ denotes the number of reactions and $\mathcal{M}$ denotes the number of protein species in the model.
The quantity $r_{j}\left(\mathbf{x},\mathbf{\epsilon},\mathbf{k}\right)$ denotes the rate of reaction $j$.
Typically, reaction $j$ is a non-linear function of biochemical and enzyme species abundance, as well as unknown model parameters $\mathbf{k}$ ($\mathcal{K}\times{1}$).
The quantity $\sigma_{ij}$ denotes the stoichiometric coefficient for species $i$ in reaction $j$.
If $\sigma_{ij}>0$, species $i$ is produced by reaction $j$.
Conversely, if $\sigma_{ij}<0$, species $i$ is consumed by reaction $j$, while $\sigma_{ij} = 0$ indicates species $i$ is not connected with reaction $j$.
Species balances were subject to the initial conditions $\mathbf{x}\left(t_{o}\right) = \mathbf{x}_{o}$.
